---
title: "Lab1_Team_Netherlands"
author: "Austin Hopkins, Ada Fitz Axen, Sarah Gaulke, Ashley Miller"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Just testing something

```{r}
#Problem 1
library(tidyr)
library(tidyselect)
library(tidyverse)
library(tinytex)

hourglass=30
#vecotr to hold pop data
x=numeric(hourglass)

#need a vector for time
time=numeric(hourglass)

#starting variable conditions
x[1]=.01
time[1]=1

#Graphics
par(mfrow=c(4,4))
par(ask=FALSE)


#Loop control
l=seq(from=.25, to=4, by=.25)
xt=function(lambda,x,t) x[t]=lambda*x[t-1]*(1-x[t-1])

m=as_tibble(matrix(nrow=length(l)*hourglass,ncol=3))
names(m)=c("Time","x","lambda")

counter = 1
for(lambda in l){
    for(t in 2:hourglass){
        x[t] = xt(lambda=lambda, x = x, t=t)
        time[t] = t
        m[counter,1:3] = t(c(t, x[t], lambda)) 
        counter = counter + 1
    }}


#plotting
#ggplot forever

m %>% filter(!is.na(x)) %>% ggplot() + aes(x = Time, y = x) + geom_line() + facet_wrap(~ lambda, scales = "free_y", labeller = label_both)  



```


##Answer problem 1 
What can you conclude about the effect of λ
 on the dynamics? Where λ is the per capita rate of population growth. 

The effect of λ is significant. If the lambda is sufficiently large it will lead to non random chaos within the population. Having a small Lambda leads to a steep decline in population. Conversely a large lamba will cause severe rise and fall of xt. 



#Problem 2


```{r}
#Problem 2

#lambda now varies in steps of 0.05 from 1 to 4
l = seq(from = 1, to = 4, by = .05)
#now 100 timesteps, not 30
tend = 100
M = as_tibble(matrix(nrow = length(l) * tend, ncol = 3))
names(M) = c("Time", "x", "lambda")
counter = 1
for(lambda in l){
for(t in 2:tend){
x[t] = xt(lambda=lambda, x=x, t=t)
time[t] = t
M[counter,1:3] = t(c(t, x[t],lambda)) #this transforms to row vector
counter = counter + 1
}
}
M <- as.data.frame(M)
M <- M[M$Time >= 51,] #removes all rows where time is less than or equal to 50
plot(M$lambda, M$x,
main = "Bifurcation Diagram",
xlab = expression(lambda),
ylab = expression(x[t]))

```
 

#Answer Problem 2
Interpretation: When λ (per capita rate of population growth) < 3, the population size x changes steadily
over time t. Once λ is equal to 3, the population over time changes rapidly, causing the population at time t
(xt) to fluctuate. Population fluctuations are at fairly regular intervals until λ>3.5, when population growth
becomes unpredictable. This tells me ecologically there is high competition when λ~3, and a population may
exceed carrying capacity once λ=4.
